\singlespacing \vspace{-1.5cm} \LARGE Trade-offs among plant floral and reproductive traits determine interactions with floral visitors

### 2021-08-16 Manuscript/Manuscript.tex
edits

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[
  12pt,
  a4paper,
]{article}
\usepackage[]{mathpazo}
\usepackage{setspace}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={ Trade-offs among plant floral and reproductive traits determine interactions with floral visitors},
  colorlinks=true,
  linkcolor=RoyalBlue,
  filecolor=Maroon,
  citecolor=Blue,
  urlcolor=RoyalBlue,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\pagestyle{plain}
\usepackage{lineno} % add 
\linenumbers % turns line numbering on

@@

\setstretch{2}
\vspace{-2.2cm}
\singlespacing
\large

\textbf{Jose B. Lanuza$^{1,2}*$, Romina Rader$^{1}$, Jamie Stavert$^{3}$, Liam K. Kendall$^{4}$, Manu E. Saunders$^{1}$ and Ignasi Bartomeus$^{2}$}

@@

\noindent

\rule{\textwidth}{1pt}

@@

There is an astonishing diversity of floral structures and plant reproductive strategies among flowering plants,\textsuperscript{\protect\hyperlink{ref-barrett2002}{1},\protect\hyperlink{ref-schiestl2013}{2}} which have long been of interest to pollination biologists in terms of their relevance to plant-pollinator interactions. However, most studies that have explored reproductive (e.g., mating and compatibility systems) and floral trait (e.g., flower size or nectar provision) variation have concentrated on the individual or community level and thus, broader macroecological patterns remain poorly investigated.\textsuperscript{\protect\hyperlink{ref-carvalheiro2014}{3}--\protect\hyperlink{ref-grossenbacher2017}{7}} Indeed, studies depicting species' life history strategies generally focus on vegetative traits and rarely consider reproductive traits.\textsuperscript{\protect\hyperlink{ref-salguero2016}{8}} As a consequence, a unified framework that explores the compromises among floral traits and their relevance to plant life strategies is currently lacking.\textsuperscript{\protect\hyperlink{ref-roddy2021}{10}} At the same time, there is growing interest in the determinants of plant-pollinator interactions via trait-based approaches\textsuperscript{\protect\hyperlink{ref-fenster2004}{11}} and trait-matching analyses.\textsuperscript{\protect\hyperlink{ref-bartomeus2016}{12}} However, floral traits have been overlooked beyond highly specialised plant-pollinator systems\textsuperscript{\protect\hyperlink{ref-roddy2021}{10},\protect\hyperlink{ref-dellinger2020}{13}} and the role of plant reproductive biology remains little explored in plant-pollinator interactions (but see references\textsuperscript{\protect\hyperlink{ref-tur2013}{14},\protect\hyperlink{ref-devaux2014}{15}}).

With the recent availability of large trait databases, plant ecological strategies are increasingly being examined,\textsuperscript{\protect\hyperlink{ref-kattge2011}{16},\protect\hyperlink{ref-salguero2015}{17}} and are facilitating the identification of global patterns and constraints of plant form and function.\textsuperscript{\protect\hyperlink{ref-salguero2016}{8},\protect\hyperlink{ref-diaz2016}{18},\protect\hyperlink{ref-carmona2021}{19}} However, the main focus has been on vegetative traits such as leaf\textsuperscript{\protect\hyperlink{ref-wright2004}{20}} or wood\textsuperscript{\protect\hyperlink{ref-chave2009}{21}} trade-offs with little or no attention given to reproductive and floral traits,\textsuperscript{\protect\hyperlink{ref-evojtko2020}{22}} also critical to plant form and function. For instance, short lived versus perennial species tend to have low versus high levels of outcrossing, respectively.\textsuperscript{\protect\hyperlink{ref-moeller2017}{6},\protect\hyperlink{ref-barrett2003}{23}} Further, outcrossing levels are positively correlated with flower size.\textsuperscript{\protect\hyperlink{ref-goodwillie2010}{24}} In addition, the presence of costly rewards (e.g., pollen or nectar) and showy flowers or floral displays can only be understood through consideration of plant species' reliance upon animal pollination (pollinator dependence) and their role in attracting pollinators.\textsuperscript{\protect\hyperlink{ref-ollerton2011}{25}} Hence, exploring plant life strategies with reproductive and floral trade-offs, in conjunction with their pollinator dependence, is necessary for a balanced understanding of plant economics.

Several studies have identified links between plant traits and plant-pollinator network properties.\textsuperscript{\protect\hyperlink{ref-carvalheiro2014}{3},\protect\hyperlink{ref-lazaro2008}{26},\protect\hyperlink{ref-bartomeus2013}{27}} Moreover, plant traits can also define species' network roles (e.g., specialists vs generalists). For example, species that occupy trait space extremes are more likely to exhibit higher specialization and be more reliant on trait-matching.\textsuperscript{\protect\hyperlink{ref-junker2013}{28},\protect\hyperlink{ref-coux2016}{29}} This morphological matching between plant and floral visitors can determine plant-pollinator interactions, and thus shape their interaction network structure.\textsuperscript{\protect\hyperlink{ref-stang2009}{30},\protect\hyperlink{ref-ibanez2012}{31}} Despite the increasing knowledge of the relevance of traits on the species network roles, little is known about how plant reproductive and floral traits determine plant species' network roles at a macroecological scale.

@@

\hypertarget{results}{%
\section{RESULTS}\label{results}}

\textbf{Plant strategies.} The phylogenetically informed principal component analysis (pPCA) captured by the first two and three axes 51.8\% and 70.97\% of trait variation, respectively (Fig. 1 and Supplementary Fig. S5) and had a phylogenetic correlation (\(\lambda\)) of 0.76. The first principal component (PC1) represented 26.72\% of the trait variation and indicated a trade-off between flower number and flower size. We refer to this axis as the `flower number - flower size trade-off' as already described in previous studies.\textsuperscript{\protect\hyperlink{ref-sargent2007}{32},\protect\hyperlink{ref-kettle2011}{33}} Hence, one end of the spectrum comprised species with high investment in flower number and plant height but small flower size, short style length and low ovule number. The other end of this spectrum comprised species that were short in height and invested in large flowers, long styles, many ovules, but few flowers. The main contributing traits to PC1 were plant height, flower number, ovule number and flower size (loadings \textgreater{} \textbar0.5\textbar; Supplementary Table S3) but style length also contributed moderately on PC1 (loading = -0.33). The second principal component (PC2) represented 25.05\% of the trait variation and indicated a trade-off between low and high pollinator dependence. We refer to this axis as the `pollinator dependence trade-off'. The main driver of trait variation on PC2 was autonomous selfing (loading = 0.85) but the other traits (except ovule number) also made moderate contributions (loadings from 0.27 to 0.4; Supplementary Table S3). We found that high pollinator dependence was associated with larger and a higher number of flowers, greater plant height and longer styles. In contrast, species with high levels of autonomous selfing tended to have fewer and smaller flowers, had shorter styles and were shorter in height. Further, PC3 explained a considerable amount of trait variability (19.17\%) and the main contributors to this axis were style length (loading = -0.66) and the degree of autonomous selfing (loading = -0.51). The remaining traits, apart from ovule number, were moderately correlated to changes on PC3 (loadings from -0.23 to -0.46; Supplementary Table S3). Thus, because style length was correlated with all traits on PC3 and was the main driver of trait variation, we refer to this axis as the `style length trade-off'. Further, the pPCA with the subset of species that had nectar and pollen quantity data showed that nectar quantity (microlitres of nectar per flower) was positively associated with flower size, style length and ovule number (PC1, 23.40\%); and pollen quantity (pollen grains per flower) was positively correlated with flower number and plant height and negatively associated with autonomous selfing (PC2, 21.67\%; Supplementary Fig. S6). This pPCA explained similar variance with the first two principal components (45.07\%) and similar associations of traits despite some variability in the loadings (Supplementary Table S4).

@@

\hypertarget{discussion}{%
\section{DISCUSSION}\label{discussion}}

@@

Over 50\% of plant trait variation was captured by the flower number - flower size and pollinator dependence trade-offs. Variation in these two axes of trait variation appears to be associated with the commonly described `fast-slow continuum' in plant\textsuperscript{\protect\hyperlink{ref-salguero2016}{8}} and animal\textsuperscript{\protect\hyperlink{ref-healy2019}{34}} life-history strategies as indicated by the different floral and reproductive traits associated with plant height, life form and lifespan. The `slow' part of the continuum (i.e., tall trees and shrubs) had plant species with many flowers, few ovules, higher pollinator dependence, greater occurrence of self-incompatibility and more complex breeding systems (e.g., monoecious and dioecious species). In contrast, species that employed the `fast' strategy (i.e., short herbs), had fewer flowers, more ovules, greater occurrence of self compatibility and lower pollinator dependence. Further, on the first two axes of trait variation, we found additional support for the positive association already described in a previous study between outcrossing rate and floral display.\textsuperscript{\protect\hyperlink{ref-goodwillie2010}{24}} This trend was further confirmed when looking at the specific correlation between floral display and autonomous selfing for our dataset (see Supplementary Fig. S10).

Despite the low predictive power of the main axes of trait variation on overall visitation patterns, we found commonalities among and within floral visitors guilds across these axes that indicate the influence of life-history strategies on floral visitation rate. For example, all guilds showed an increasing trend of visitation on plant species with higher pollinator dependence, which were also associated with larger floral displays and greater pollen quantities (Fig. 1 and Supplementary Fig. S6). This trend is consistent with previous plant-pollinator studies that show higher visitation rates on species that make greater reproductive investment.\textsuperscript{\protect\hyperlink{ref-hegland2005}{35}--\protect\hyperlink{ref-kaiser2014}{37}} In regard to the flower number - flower size and style length trade-offs, different guilds showed contrasting visitation rates, which could be associated with different pollination syndromes at a macroecological scale. For instance, the guilds of bees and syrphids were clearly associated with opposing life-strategies on PC1 and PC3 (Fig. 3) indicating a possible niche partitioning among these two guilds.\textsuperscript{\protect\hyperlink{ref-palmer2003}{38},\protect\hyperlink{ref-phillips2020}{39}} However, these plant-pollinator associations do not account for some of the traits (i.e., pollen and nectar quantity) that had greater relevance on the species network roles (Fig. 4) because of insufficient data availability and lack of pollinator efficiency measurements for more complete descriptions of the pollination syndromes.\textsuperscript{\protect\hyperlink{ref-dellinger2020}{13}} In any case, it is worth noting that this general pattern emerges in field-measured plant-pollinator networks, where other local factors such as species relative abundances, will surely explain part of the observed variability.\textsuperscript{\protect\hyperlink{ref-bartomeus2016}{12},\protect\hyperlink{ref-vazquez2007}{40},\protect\hyperlink{ref-encinas2012}{41}}

To conclude, we provide the first description of plant floral and reproductive trade-offs using a large global dataset of plant traits and interactions with floral visitors. This allowed us to identify the major reproductive strategies of flowering plants and how plant life strategies determine their interaction with floral visitor guilds. Although the percentage of explained variation found with the first two axes (\textasciitilde50\%) is lower than other studies that have evaluated life histories with vegetative traits,\textsuperscript{\protect\hyperlink{ref-diaz2016}{18},\protect\hyperlink{ref-carmona2021}{19}} this percentage is consistent with the largest and most recent study that have characterized life strategies with both vegetative and reproductive (i.e., semelparity and iteroparity) traits.\textsuperscript{\protect\hyperlink{ref-salguero2016}{8}} Future work needs to integrate these floral and reproductive compromises with the already described vegetative and physiological trade-offs\textsuperscript{\protect\hyperlink{ref-diaz2016}{18}} to create a more comprehensive spectra of trait variation.\textsuperscript{\protect\hyperlink{ref-roddy2021}{10}} Further, the varying level of phylogenetic signal among traits deserves further attention to understand evolutionary changes on mating and flower morphology in response to pollinators.\textsuperscript{\protect\hyperlink{ref-gervasi2017}{42},\protect\hyperlink{ref-mackin2021}{43}} Finally, a more complete description of the macroecological patterns of plant-pollinator interactions including unrepresented areas of the world\textsuperscript{\protect\hyperlink{ref-poisot2021}{44}} and traits that have been largely ignored (e.g., floral rewards and plant mating system) are needed to progress in our understanding of the global patterns in plant-pollinator interactions.

\hypertarget{materials-and-methods}{%
\section{MATERIALS AND METHODS}\label{materials-and-methods}}

\textbf{Plant-pollinator network studies.} We selected 28 studies from 18 different countries that constituted a total of 64 plant-pollinator networks. These studies recorded plant-pollinator interactions in natural systems and were selected so that we had broad geographical representation. Although these studies differ in sampling effort and methodology, all studies provided information about plant-pollinator interactions (weighted and non-weighted), which we used to build a database of plant species that are likely to be animal pollinated. Many of these networks are freely available either as published studies\textsuperscript{\protect\hyperlink{ref-carvalheiro2014}{3},\protect\hyperlink{ref-fortuna2010}{45}} or available in online archives (e.g., `The Web of Life'\textsuperscript{\protect\hyperlink{ref-fortuna2010}{45}} and `Mangal'.\textsuperscript{\protect\hyperlink{ref-poisot2016}{46}} In total, our network dataset (see Supplementary Table S1) constituted 60 weighted (interaction frequency) and 4 unweighted (presence/absence of the interaction) networks, each sampled at a unique location and year, as well as eight meta-webs where interactions were pooled across several locations and multiple years.

\textbf{Taxonomy of plants and pollinators.} All species names, genera, families and orders were retrieved and standardized from the taxonomy data sources NCBI (\url{https://www.ncbi.nlm.nih.gov/taxonomy}) for plants and ITIS (\url{https://www.itis.gov/}) for pollinators, using the R package \emph{taxize}.\textsuperscript{\protect\hyperlink{ref-chamberlain2020}{47}} We filled the `not found' searches manually using \url{http://www.theplantlist.org/} and \url{http://www.mobot.org/} for plants and \url{http://www.catalogueoflife.org/} for floral visitors.

\textbf{Functional traits.} We selected 20 different functional traits based on their relevance to plant reproduction and data availability (Table 1). These included twelve quantitative and eight categorical traits belonging to three broader trait groupings (13 floral, 4 reproductive and 3 vegetative). For each plant species, we undertook an extensive literature and online search across a wide range of resources (plant databases, online floras, books, journals and images). From a total of 30,120 cells (20 columns × 1,506 species) we were able to fill 24,341 cells (80.8\% of the dataset, see Supplementary Fig. S1 for missing values information for each trait).

\textbf{Phylogenetic Distance.} We calculated the phylogenetic distance between different plant species using the function \emph{get\_tree} from the package \emph{rtrees} (\url{https://github.com/daijiang/rtrees}), which downloads phylogenetic distances from the extended R implementation of the Open Tree of Life.\textsuperscript{\protect\hyperlink{ref-smith2018}{48},\protect\hyperlink{ref-jin2019}{49}}

@@

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-barrett2002}{}%
1. Barrett, S. C. H. The evolution of plant sexual diversity. \emph{Nat. Rev. Genet.} \textbf{3}, 274--284 (2002).

>>>>

\documentclass[12pt,a4paper,]{article}
\usepackage[]{mathpazo}
\usepackage{setspace}
\setstretch{2}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\PassOptionsToPackage{hyphens}{url} % url is loaded by hyperref
\usepackage[unicode=true]{hyperref}
\PassOptionsToPackage{usenames,dvipsnames}{color} % color is loaded by hyperref
\hypersetup{
            pdftitle={ Trade-offs among plant floral and reproductive traits determine interactions with floral visitors},
            colorlinks=true,
            linkcolor=RoyalBlue,
            citecolor=Blue,
            urlcolor=RoyalBlue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage[margin=1in]{geometry}
\usepackage{longtable,booktabs}
% Fix footnotes in tables (requires footnote package)
\IfFileExists{footnote.sty}{\usepackage{footnote}\makesavenoteenv{long table}}{}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\usepackage{lineno} % add 
\linenumbers % turns line numbering on

@@

\vspace{-2.2cm} \singlespacing
\large 
\textbf{Jose B. Lanuza$^{1,2}*$, Romina Rader$^{1}$, Jamie Stavert$^{3}$, Liam K. Kendall$^{4}$, Manu E. Saunders$^{1}$ and Ignasi Bartomeus$^{2}$}

@@

\noindent\rule{\textwidth}{1pt}

@@

There is an astonishing diversity of floral structures and plant
reproductive strategies among flowering
plants\textsuperscript{\protect\hyperlink{ref-barrett2002}{1},\protect\hyperlink{ref-schiestl2013}{2}},
which have long been of interest to pollination biologists in terms of
their relevance to plant-pollinator interactions. However, most studies
that have explored reproductive (e.g., mating and compatibility systems)
and floral trait (e.g., flower size or nectar provision) variation have
concentrated on the individual or community level and thus, broader
macroecological patterns remain poorly
investigated\textsuperscript{\protect\hyperlink{ref-carvalheiro2014}{3}--\protect\hyperlink{ref-grossenbacher2017}{7}}.
Indeed, studies depicting species' life history strategies generally
focus on vegetative traits and rarely consider reproductive
traits\textsuperscript{\protect\hyperlink{ref-salguero2016}{8}}. As a
consequence, a unified framework that explores the compromises among
floral traits and their relevance to plant life strategies is currently
lacking\textsuperscript{\protect\hyperlink{ref-roddy2021}{10}}. At the
same time, there is growing interest in the determinants of
plant-pollinator interactions via trait-based
approaches\textsuperscript{\protect\hyperlink{ref-fenster2004}{11}} and
trait-matching
analyses\textsuperscript{\protect\hyperlink{ref-bartomeus2016}{12}}.
However, floral traits have been overlooked beyond highly specialised
plant-pollinator
systems\textsuperscript{\protect\hyperlink{ref-roddy2021}{10},\protect\hyperlink{ref-dellinger2020}{13}}
and the role of plant reproductive biology remains little explored in
plant-pollinator interactions (but see
references\textsuperscript{\protect\hyperlink{ref-tur2013}{14},\protect\hyperlink{ref-devaux2014}{15}}).

With the recent availability of large trait databases, plant ecological
strategies are increasingly being
examined\textsuperscript{\protect\hyperlink{ref-kattge2011}{16},\protect\hyperlink{ref-salguero2015}{17}},
and are facilitating the identification of global patterns and
constraints of plant form and
function\textsuperscript{\protect\hyperlink{ref-salguero2016}{8},\protect\hyperlink{ref-diaz2016}{18},\protect\hyperlink{ref-carmona2021}{19}}.
However, the main focus has been on vegetative traits such as
leaf\textsuperscript{\protect\hyperlink{ref-wright2004}{20}} or
wood\textsuperscript{\protect\hyperlink{ref-chave2009}{21}} trade-offs
with little or no attention given to reproductive and floral
traits\textsuperscript{\protect\hyperlink{ref-evojtko2020}{22}}, also
critical to plant form and function. For instance, short lived versus
perennial species tend to have low versus high levels of outcrossing,
respectively\textsuperscript{\protect\hyperlink{ref-moeller2017}{6},\protect\hyperlink{ref-barrett2003}{23}}.
Further, outcrossing levels are positively correlated with flower
size\textsuperscript{\protect\hyperlink{ref-goodwillie2010}{24}}. In
addition, the presence of costly rewards (e.g., pollen or nectar) and
showy flowers or floral displays can only be understood through
consideration of plant species' reliance upon animal pollination
(pollinator dependence) and their role in attracting
pollinators\textsuperscript{\protect\hyperlink{ref-ollerton2011}{25}}.
Hence, exploring plant life strategies with reproductive and floral
trade-offs, in conjunction with their pollinator dependence, is
necessary for a balanced understanding of plant economics.

Several studies have identified links between plant traits and
plant-pollinator network
properties\textsuperscript{\protect\hyperlink{ref-carvalheiro2014}{3},\protect\hyperlink{ref-lazaro2008}{26},\protect\hyperlink{ref-bartomeus2013}{27}}.
Moreover, plant traits can also define species' network roles (e.g.,
specialists vs generalists). For example, species that occupy trait
space extremes are more likely to exhibit higher specialization and be
more reliant on
trait-matching\textsuperscript{\protect\hyperlink{ref-junker2013}{28},\protect\hyperlink{ref-coux2016}{29}}.
This morphological matching between plant and floral visitors can
determine plant-pollinator interactions, and thus shape their
interaction network
structure\textsuperscript{\protect\hyperlink{ref-stang2009}{30},\protect\hyperlink{ref-ibanez2012}{31}}.
Despite the increasing knowledge of the relevance of traits on the
species network roles, little is known about how plant reproductive and
floral traits determine plant species' network roles at a
macroecological scale.

@@

\section{RESULTS}\label{results}

\textbf{Plant strategies.} The phylogenetically informed principal
component analysis (pPCA) captured by the first two and three axes
51.8\% and 70.97\% of trait variation, respectively (Fig. 1 and
Supplementary Fig. S5) and had a phylogenetic correlation (\(\lambda\))
of 0.76. The first principal component (PC1) represented 26.72\% of the
trait variation and indicated a trade-off between flower number and
flower size. We refer to this axis as the `flower number - flower size
trade-off' as already described in previous
studies\textsuperscript{\protect\hyperlink{ref-sargent2007}{32},\protect\hyperlink{ref-kettle2011}{33}}.
Hence, one end of the spectrum comprised species with high investment in
flower number and plant height but small flower size, short style length
and low ovule number. The other end of this spectrum comprised species
that were short in height and invested in large flowers, long styles,
many ovules, but few flowers. The main contributing traits to PC1 were
plant height, flower number, ovule number and flower size (loadings
\textgreater{} \textbar{}0.5\textbar{}; Supplementary Table S3) but
style length also contributed moderately on PC1 (loading = -0.33). The
second principal component (PC2) represented 25.05\% of the trait
variation and indicated a trade-off between low and high pollinator
dependence. We refer to this axis as the `pollinator dependence
trade-off'. The main driver of trait variation on PC2 was autonomous
selfing (loading = 0.85) but the other traits (except ovule number) also
made moderate contributions (loadings from 0.27 to 0.4; Supplementary
Table S3). We found that high pollinator dependence was associated with
larger and a higher number of flowers, greater plant height and longer
styles. In contrast, species with high levels of autonomous selfing
tended to have fewer and smaller flowers, had shorter styles and were
shorter in height. Further, PC3 explained a considerable amount of trait
variability (19.17\%) and the main contributors to this axis were style
length (loading = -0.66) and the degree of autonomous selfing (loading =
-0.51). The remaining traits, apart from ovule number, were moderately
correlated to changes on PC3 (loadings from -0.23 to -0.46;
Supplementary Table S3). Thus, because style length was correlated with
all traits on PC3 and was the main driver of trait variation, we refer
to this axis as the `style length trade-off'. Further, the pPCA with the
subset of species that had nectar and pollen quantity data showed that
nectar quantity (microlitres of nectar per flower) was positively
associated with flower size, style length and ovule number (PC1,
23.40\%); and pollen quantity (pollen grains per flower) was positively
correlated with flower number and plant height and negatively associated
with autonomous selfing (PC2, 21.67\%; Supplementary Fig. S6). This pPCA
explained similar variance with the first two principal components
(45.07\%) and similar associations of traits despite some variability in
the loadings (Supplementary Table S4).

@@

\section{DISCUSSION}\label{discussion}

@@

Over 50\% of plant trait variation was captured by the flower number -
flower size and pollinator dependence trade-offs. Variation in these two
axes of trait variation appears to be associated with the commonly
described `fast-slow continuum' in
plant\textsuperscript{\protect\hyperlink{ref-salguero2016}{8}} and
animal\textsuperscript{\protect\hyperlink{ref-healy2019}{34}}
life-history strategies as indicated by the different floral and
reproductive traits associated with plant height, life form and
lifespan. The `slow' part of the continuum (i.e., tall trees and shrubs)
had plant species with many flowers, few ovules, higher pollinator
dependence, greater occurrence of self-incompatibility and more complex
breeding systems (e.g., monoecious and dioecious species). In contrast,
species that employed the `fast' strategy (i.e., short herbs), had fewer
flowers, more ovules, greater occurrence of self compatibility and lower
pollinator dependence. Further, on the first two axes of trait
variation, we found additional support for the positive association
already described in a previous study between outcrossing rate and
floral
display\textsuperscript{\protect\hyperlink{ref-goodwillie2010}{24}}.
This trend was further confirmed when looking at the specific
correlation between floral display and autonomous selfing for our
dataset (see Supplementary Fig. S10).

Despite the low predictive power of the main axes of trait variation on
overall visitation patterns, we found commonalities among and within
floral visitors guilds across these axes that indicate the influence of
life-history strategies on floral visitation rate. For example, all
guilds showed an increasing trend of visitation on plant species with
higher pollinator dependence, which were also associated with larger
floral displays and greater pollen quantities (Fig. 1 and Supplementary
Fig. S6). This trend is consistent with previous plant-pollinator
studies that show higher visitation rates on species that make greater
reproductive
investment\textsuperscript{\protect\hyperlink{ref-hegland2005}{35}--\protect\hyperlink{ref-kaiser2014}{37}}.
In regard to the flower number - flower size and style length
trade-offs, different guilds showed contrasting visitation rates, which
could be associated with different pollination syndromes at a
macroecological scale. For instance, the guilds of bees and syrphids
were clearly associated with opposing life-strategies on PC1 and PC3
(Fig. 3) indicating a possible niche partitioning among these two
guilds\textsuperscript{\protect\hyperlink{ref-palmer2003}{38},\protect\hyperlink{ref-phillips2020}{39}}.
However, these plant-pollinator associations do not account for some of
the traits (i.e., pollen and nectar quantity) that had greater relevance
on the species network roles (Fig. 4) because of insufficient data
availability and lack of pollinator efficiency measurements for more
complete descriptions of the pollination
syndromes\textsuperscript{\protect\hyperlink{ref-dellinger2020}{13}}. In
any case, it is worth noting that this general pattern emerges in
field-measured plant-pollinator networks, where other local factors such
as species relative abundances, will surely explain part of the observed
variability\textsuperscript{\protect\hyperlink{ref-bartomeus2016}{12},\protect\hyperlink{ref-vazquez2007}{40},\protect\hyperlink{ref-encinas2012}{41}}.

To conclude, we provide the first description of plant floral and
reproductive trade-offs using a large global dataset of plant traits and
interactions with floral visitors. This allowed us to identify the major
reproductive strategies of flowering plants and how plant life
strategies determine their interaction with floral visitor guilds.
Although the percentage of explained variation found with the first two
axes (\textasciitilde{}50\%) is lower than other studies that have
evaluated life histories with vegetative
traits\textsuperscript{\protect\hyperlink{ref-diaz2016}{18},\protect\hyperlink{ref-carmona2021}{19}},
this percentage is consistent with the largest and most recent study
that have characterized life strategies with both vegetative and
reproductive (i.e., semelparity and iteroparity)
traits\textsuperscript{\protect\hyperlink{ref-salguero2016}{8}}. Future
work needs to integrate these floral and reproductive compromises with
the already described vegetative and physiological
trade-offs\textsuperscript{\protect\hyperlink{ref-diaz2016}{18}} to
create a more comprehensive spectra of trait
variation\textsuperscript{\protect\hyperlink{ref-roddy2021}{10}}.
Further, the varying level of phylogenetic signal among traits deserves
further attention to understand evolutionary changes on mating and
flower morphology in response to
pollinators\textsuperscript{\protect\hyperlink{ref-gervasi2017}{42},\protect\hyperlink{ref-mackin2021}{43}}.
Finally, a more complete description of the macroecological patterns of
plant-pollinator interactions including unrepresented areas of the
world\textsuperscript{\protect\hyperlink{ref-poisot2021}{44}} and traits
that have been largely ignored (e.g., floral rewards and plant mating
system) are needed to progress in our understanding of the global
patterns in plant-pollinator interactions.

\section{MATERIALS AND METHODS}\label{materials-and-methods}

\textbf{Plant-pollinator network studies.} We selected 28 studies from
18 different countries that constituted a total of 64 plant-pollinator
networks. These studies recorded plant-pollinator interactions in
natural systems and were selected so that we had broad geographical
representation. Although these studies differ in sampling effort and
methodology, all studies provided information about plant-pollinator
interactions (weighted and non-weighted), which we used to build a
database of plant species that are likely to be animal pollinated. Many
of these networks are freely available either as published
studies\textsuperscript{\protect\hyperlink{ref-carvalheiro2014}{3},\protect\hyperlink{ref-fortuna2010}{45}}
or available in online archives (e.g., `The Web of
Life'\textsuperscript{\protect\hyperlink{ref-fortuna2010}{45}} and
`Mangal'\textsuperscript{\protect\hyperlink{ref-poisot2016}{46}}. In
total, our network dataset (see Supplementary Table S1) constituted 60
weighted (interaction frequency) and 4 unweighted (presence/absence of
the interaction) networks, each sampled at a unique location and year,
as well as eight meta-webs where interactions were pooled across several
locations and multiple years.

\textbf{Taxonomy of plants and pollinators.} All species names, genera,
families and orders were retrieved and standardized from the taxonomy
data sources NCBI (\url{https://www.ncbi.nlm.nih.gov/taxonomy}) for
plants and ITIS (\url{https://www.itis.gov/}) for pollinators, using the
R package
\emph{taxize}\textsuperscript{\protect\hyperlink{ref-chamberlain2020}{47}}.
We filled the `not found' searches manually using
\url{http://www.theplantlist.org/} and \url{http://www.mobot.org/} for
plants and \url{http://www.catalogueoflife.org/} for floral visitors.

\textbf{Functional traits.} We selected 20 different functional traits
based on their relevance to plant reproduction and data availability
(Table 1). These included twelve quantitative and eight categorical
traits belonging to three broader trait groupings (13 floral, 4
reproductive and 3 vegetative, Supplementary Information). For each
plant species, we undertook an extensive literature and online search
across a wide range of resources (plant databases, online floras, books,
journals and images). From a total of 30,120 cells (20 columns × 1,506
species) we were able to fill 24,341 cells (80.8\% of the dataset, see
Supplementary Fig. S1 for missing values information for each trait).

\textbf{Phylogenetic Distance.} We calculated the phylogenetic distance
between different plant species using the function \emph{get\_tree} from
the package \emph{rtrees} (\url{https://github.com/daijiang/rtrees}),
which downloads phylogenetic distances from the extended R
implementation of the Open Tree of
Life\textsuperscript{\protect\hyperlink{ref-smith2018}{48},\protect\hyperlink{ref-jin2019}{49}}.

@@

\section{References}\label{references}

\hypertarget{refs}{}
\hypertarget{ref-barrett2002}{}
1. Barrett, S. C. H. The evolution of plant sexual diversity. \emph{Nat.
Rev. Genet.} \textbf{3}, 274--284 (2002).

\hypertarget{ref-schiestl2013}{}
2. Schiestl, F. P. \& Johnson, S. D. Pollinator-mediated evolution of
floral signals. \emph{Trends Ecol. Evol.} \textbf{28}, 307--315 (2013).
